---
title: "Compute ROI-level ISC from Haemo data"
---

Setup and define parameters
```{r}
library(tidyverse)

base_dir <- "/Users/saewonchung/Desktop/fNIRS pilot analysis/Data_preprocessed"
preproc_csv_filepattern <- "sub-\\d+_task-Video_acq-\\d+_dur-508_desc-preproc_haemo\\.csv$"

# map ROI to list of channels
roi_map <- list(
    vmPFC = c("S1_D1", "S1_D9", "S1_D8", "S2_D9", "S10_D9"),
    dmPFC = c("S9_D9", "S9_D3", "S9_D11", "S9_D4", "S5_D3", "S5_D4", "S13_D4", "S13_D11"),
    ldlPFC = c("S5_D5", "S4_D5", "S4_D3", "S2_D3", "S2_D2", "S3_D2"),
    rdlPFC = c("S13_D12", "S12_D12", "S12_D11", "S10_11", "S10_D10", "S11_D10"),
    lTPJ = c("S8_D7", "S8_D6", "S7_D7", "S7_D6"),
    rTPJ = c("S15_D14", "S15_D13", "S16_D14", "S16_D13")
)
```

Read the preprocessed HbO/HbR timeseries
```{r}
file_paths <- list.files(path = base_dir,
                         pattern = preproc_csv_filepattern,
                         recursive = TRUE,
                         full.names = TRUE)

subject_ids <- str_extract(file_paths, "(?<=sub-)\\d+")
ordering <- order(as.numeric(subject_ids))
file_paths <- file_paths[ordering]
subject_ids <- subject_ids[ordering]

subject_data <- lapply(file_paths, function(f) {
  read_csv(f, show_col_types = FALSE) %>% select(-time)
})

names(subject_data) <- subject_ids
```

Trim to shortest length
```{r}
lengths_check <- sapply(subject_data, nrow) # length check
print(lengths_check)
min_len <- min(sapply(subject_data, nrow)) # if length different,
subject_data <- lapply(subject_data, function(df) df[1:min_len, ]) # trim to min length
```

Z score within participant within channel
```{r}
subject_data <- lapply(subject_data, function(df) {
  df %>% mutate(across(everything(), scale))
}) 
```

Average by ROI
```{r}
average_by_roi <- function(subject_data, signal_type = c("hbo", "hbr")) {
  signal_type <- match.arg(signal_type)
  roi_map_hbx <- lapply(roi_map, function(chs) paste(chs, signal_type))
  
  lapply(subject_data, function(df) {
    df_filtered <- df %>% select(matches(signal_type))
    
    roi_data <- map_dfc(roi_map_hbx, function(chs) {
      valid_chs <- intersect(chs, colnames(df_filtered))
      if (length(valid_chs) > 0) {
        rowMeans(df_filtered[, valid_chs, drop = FALSE], na.rm = TRUE)
      } else {
        rep(NA, nrow(df_filtered))
      }
    })
    
    colnames(roi_data) <- names(roi_map_hbx)
    as_tibble(roi_data)
  })
}

subject_data_hbo_roi <- average_by_roi(subject_data, "hbo")
subject_data_hbr_roi <- average_by_roi(subject_data, "hbr")
```

Compute Dyad-Level ISC per Channel (applying Fisher's z-transformation to the resulting ISC values)
```{r}
fisher_z <- function(r) 0.5 * log((1 + r) / (1 - r), base=exp(1)) # equivalent to atanh(r)
inverse_fisher_z <- function(z) (exp(2 * z) - 1) / (exp(2 * z) + 1) # equivalent to tanh(z)

channel_names_hbo <- unique(unlist(lapply(subject_data_hbo_roi, colnames)))
channel_names_hbr <- unique(unlist(lapply(subject_data_hbr_roi, colnames)))

compute_dyad_isc <- function(subject_data_roi, channel_names) {
  dyad_isc_list <- list()
  
  for (ch_name in channel_names) {
    ts_all <- lapply(subject_data_roi, function(x) x[[ch_name]])
    available <- which(sapply(ts_all, function(x) is.numeric(x) && length(x) == length(ts_all[[1]])))
    
    if (length(available) < 2) next
    
    for (i in 1:(length(available) - 1)) {
      for (j in (i + 1):length(available)) {
        sub1 <- names(subject_data_roi)[available[i]]
        sub2 <- names(subject_data_roi)[available[j]]
        r <- cor(ts_all[[sub1]], ts_all[[sub2]], use = "complete.obs")
        z <- fisher_z(r)
        
        dyad_isc_list[[length(dyad_isc_list) + 1]] <- tibble(
          channel = ch_name,
          sub1 = sub1,
          sub2 = sub2,
          isc_z = z
        )
      }
    }
  }
}

dyad_isc_hbo <- compute_dyad_isc(subject_data_hbo_roi, channel_names_hbo)
dyad_isc_hbr <- compute_dyad_isc(subject_data_hbr_roi, channel_names_hbr)
```

Compute individual-Level ISC per Channel (by getting participant's average dyad-level ISC across the dyads they're in)
```{r}
to_individual_isc <- function(dyad_df, signal_type) {
  dyad_df %>%
    pivot_longer(cols = c(sub1, sub2), names_to = "role", values_to = "subject") %>%
    group_by(subject, channel) %>%
    summarize(mean_isc_z = mean(isc_z, na.rm = TRUE), .groups = "drop") %>%
    mutate(signal_type = signal_type)
}

roi_isc_hbo <- to_individual_isc(dyad_isc_hbo, "hbo")
roi_isc_hbr <- to_individual_isc(dyad_isc_hbr, "hbr")
```

Save results to csv
```{r}
roi_isc <- bind_rows(roi_isc_hbo, roi_isc_hbr) %>% 
  mutate(subject = as.numeric(subject))

write_csv(roi_isc, "/Users/saewonchung/Desktop/fNIRS pilot analysis/ISC_ROI_level.csv")
```
